\section{SBD Foundations: TransNetV2 and AutoShot}
\label{sec:foundation}

This section formulates the shot boundary detection problem and reviews the
architectural components of the TransNetV2 family and the AutoShot search
that produced the backbone AutoShotV2 builds upon.

\subsection{Problem Formulation}
\label{sec:problem}

Given a video represented as an ordered sequence of $T$ frames
$\{x_t\}_{t=1}^{T}$, SBD predicts a binary sequence $\{b_t\}_{t=1}^{T}$,
where $b_t{=}1$ denotes a shot boundary (an abrupt cut or a gradual
transition) at frame $t$. A model outputs a boundary probability sequence
$\{p_t\}_{t=1}^{T}$ with $p_t\in[0,1]$, and the final boundary set is
obtained by thresholding:
\begin{equation}
    \hat b_t=\begin{cases}1, & p_t > \theta,\\ 0, & p_t \le \theta,\end{cases}
    \label{eq:threshold}
\end{equation}
where $\theta$ is selected on validation data. Following standard practice,
a predicted boundary counts as correct when it falls within
\PaperMatchTolerance{} frames of a ground-truth boundary, and consecutive
positive predictions are grouped into one transition.

\subsection{Spatial-Temporal Factorized Convolution}

Processing raw frame sequences with full 3-D convolutional kernels of size
$k\times k\times k$ is expensive. The TransNetV2 family therefore factorizes
each 3-D convolution into a 2-D spatial convolution ($1\times k\times k$,
learning edges, textures, and color regions per frame) followed by a 1-D
temporal convolution ($k\times 1\times 1$, capturing motion patterns along
the temporal axis). With $C$ input/output channels and $k{=}3$, the
parameter count drops from $27C^2$ to $9C^2+3C^2=12C^2$, a reduction of
roughly 56\% that also imposes a useful inductive bias separating spatial
content from temporal motion.

\subsection{Dilated DDCNN Blocks}

The backbone stacks Dilated Convolutional Neural Network (DDCNN) blocks.
Each block splits its input features into four parallel temporal-convolution
branches with exponentially increasing dilation rates $d\in\{1,2,4,8\}$,
concatenates the branch outputs along the channel dimension, and applies
batch normalization with ReLU activation. For temporal kernel size $K{=}3$,
the cumulative receptive field after block $k$ grows as
\begin{equation}
    R_k = R_{k-1} + 2\,(K-1)\!\!\sum_{d\in\{1,2,4,8\}}\!\! d \;=\; R_{k-1} + 60,
    \label{eq:receptive}
\end{equation}
with $R_0{=}1$. Six stacked blocks yield a theoretical receptive field of up
to 361 frames; since inference uses sliding windows of 100 frames, the
effective context is capped by the window but remains wide enough for long
dissolves and fades.

\subsection{Similarity Branch}

A parallel similarity branch complements the convolutional features with two
forms of visual consistency: (i) a 512-bin RGB histogram per frame, whose
cosine distances between consecutive frames yield a 101-dimensional
color-shift descriptor, and (ii) learned similarity, the cosine similarity
between normalized deep features from intermediate convolutional layers.
Both are fused through a fully connected layer with 128 units and ReLU
activation, and the result is concatenated with the convolutional features
before classification.

\subsection{AutoShot Backbone Search}

While TransNetV2 fixes a configuration of identical DDCNN blocks, AutoShot
searches the backbone configuration for short-form video with neural
architecture search~\cite{zhu2023autoshot,elsken2019nas,white2023_nas1000,ren2020_nas_survey}.
The space covers four DDCNN block variants (differing in channel
configuration and spatial feature sharing) across six layers plus a choice
of zero to four Transformer layers~\cite{vaswani2017attention} at the
seventh, about $83.9\times10^6$ candidate architectures in total. The search
trains a single-path one-shot supernet~\cite{guo2020spos} explored with
Bayesian optimization (differentiable alternatives such as
DARTS~\cite{liu2019darts} were considered by the original authors), and the
selected model is refined with knowledge distillation and weight grafting.
The architecture found under the F1 criterion uses DDCNNV2 blocks of 4--12
filters with shared spatial convolutions in layers 2--4 and no Transformer
layer. Its penultimate stage exposes a 4864-dimensional feature vector,
which we freeze and cache:
\begin{equation}
    h_t \in \mathbb{R}^{4864}.
    \label{eq:features}
\end{equation}
All AutoShotV2 components operate on $h_t$ without updating the backbone.
