\begin{abstract}
Shot boundary detection (SBD) is a fundamental preprocessing step for video indexing, retrieval, summarization, and automated editing. However, modern short-video platforms create a difficult operating regime: videos are brief, shots are extremely short, transitions are visually complex, and boundary frames are rare compared with non-boundary frames. This paper presents AutoShotV2, a lightweight improvement over the AutoShot architecture for short-video SBD. Instead of re-searching or retraining the full backbone, AutoShotV2 freezes the AutoShot feature extractor and trains a new two-branch classification head over 4864-dimensional cached features. The head is optimized with Focal Loss and complementary one-hot and boundary-aware many-hot supervision. At inference time, the primary boundary logits are calibrated by temperature scaling and stabilized by one-dimensional Gaussian smoothing before thresholding. On the SHOT test set of 200 short videos, AutoShotV2 reaches an F1-score of \PaperASVTwoShotFOne, with precision of \PaperASVTwoShotPrecision{} and recall of \PaperASVTwoShotRecall, improving over the reproduced AutoShot baseline by \PaperASVTwoVsAutoShotPP{} F1 points and over the reported TransNetV2 baseline by \PaperASVTwoVsTransNetPP{} points. Experiments on BBC and ClipShots show that the same fixed deployment configuration transfers well to documentary footage but remains limited on open-world gradual transitions. The code is available at \url{https://github.com/DoTienManhCoder/AutoShotV2}.
\end{abstract}
